How Slow do my Transient Typechecks Go?

### 2019-08-17 paper.tex

\begin{abstract}
One form of type checking used in gradually typed language is
\emph{transient type checking}: when ever an object `flows' through
code with a type annotation, the object is dynamically checked to
ensure it has the methods required by the annotation.

@@

Grace is an object-oriented, imperative, educational programming
language, with a focus on introductory programming
courses, but also intended for more advanced study and research \citep{graceOnward12,graceSigcse13}.
%
While Grace's syntax draws
from the so-called ``curly bracket'' traditions of C, Java, and
JavaScript, the structure of the language
is in many ways closer to Smalltalk:
all computation is done via dynamically dispatched  ``method requests''
where the object receiving the request decides which
%kjx reverted cos edited version made no sense.
%for this paper, we could just delete the comments about blocks if you like.
method to run, control structures are built out of blocks (essentialy lambda expressions), and return statements are ``non-local'', returning from the lexically enclosing method \citep{bluebook}.
%
In
other ways, Grace is closer to JavaScript than Smalltalk: Grace
objects are created from object literals, rather than by
instantiating classes \citep{Black2007-emeraldHOPL,JonesECOOP2016} and
objects and classes can be deeply nested within each 
other \citep{betabook}.

@@

Though outliers remain visible in the plots for each individual benchmark, the largest 95\% confidence interval we obtained for the arithmetic mean of any of the results for any of our experiments was $\pm3.53\%$ (for the PyStone benchmark).

All experiments were executed on a Windows 10 machine
running Fedora 29 within Windows Subsystems for Linux (version 1). The machine has 
one Intel Core i7-6800K @3.40GHz, with 6 cores, for a total of 12 hyperthreads. We used ReBench 1.0 \citep{ReBench:2018}, Java 1.8.0\_191 Graal 0.43.
Benchmarks were executed one by one to avoid interference between them.
The analysis of the results and plots where generated using Python 3.7.3 and PGFPLOTS 1.16.
To enable reproductions, the scripts we used to generate and run our experiments, including the source code for all the configurations tested, are available online.\footnote{
\url{https://gitlab.ecs.vuw.ac.nz/isaac/Moth-Takikawa}}

>>>>

\begin{abstract}
One form of type checking used in gradually typed language is
\emph{transient type checking}: whenever an object `flows' through
code with a type annotation, the object is dynamically checked to
ensure it has the methods required by the annotation.

@@

Grace is an object-oriented, imperative, educational programming
language, with a focus on introductory programming
courses, but also intended for more advanced study and research \citep{graceOnward12,graceSigcse13}.
%
While Grace's syntax draws
from the so-called ``curly bracket'' traditions of C, Java, and
JavaScript, the structure of the language
is in many ways closer to Smalltalk:
all computation is done via dynamically dispatched  ``method requests''
where the object receiving the request decides what code to run,
control structures are built out of lambda expressions support ``non-local'' returns, i.e. they can return to the point where execution first encountered the lambda \citep{bluebook}.  In
other ways, Grace is closer to JavaScript than Smalltalk: Grace
objects are created from object literals, rather than by
instantiating classes \citep{Black2007-emeraldHOPL,JonesECOOP2016} and
objects and classes can be deeply nested within each 
other \citep{betabook}.

@@

Though outliers remain visible in the plots for each individual benchmark, the largest 95\% confidence interval we obtained (over the mean time after warmup) for any of experiments was $\pm3.53\%$ (for the PyStone benchmark).

All experiments were executed on a Windows 10 machine
running Fedora 30 within Windows Subsystems for Linux (version 1). The machine has 
an Intel Core i7-6800K @3.40GHz, with 6 cores, for a total of 12 hyperthreads. We used ReBench 1.0 \citep{ReBench:2018}, Java 1.8.0\_191 Graal 0.43.
Benchmarks were executed one by one to avoid interference between them.
The analysis of the results and plots where generated using Python 3.7.3 and PGFPLOTS 1.16.
To enable reproductions, the scripts we used to generate and run our experiments, including the source code for all the configurations tested, are available online.\footnote{
\url{https://gitlab.ecs.vuw.ac.nz/isaac/Moth-Takikawa}}
